---
title: "In-class Exercise 1"
author: "Ho Zi Jun"
date: "April 13, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Loading R packages

In the code chunk below, [`p_load()`](https://www.geeksforgeeks.org/p_load-function-in-r/) of [pacman](https://trinker.github.io/pacman_dev/) package is used to load the tidyverse family of packages.

```{r}
pacman::p_load(tidyverse)
```

## Comparison of functions for reading csv files.

::: panel-tabset
## `read_csv()`

```{r}
realis <- read_csv("data/realis2019.csv")
```

```{r}
class(realis)
```

For the `read_csv()` function it reads the file into a tibble object

## `read.csv()`

```{r}
realis.csv <- read.csv("data/realis2019.csv")
```

```{r}
class(realis.csv)
```

For the `read.csv()` it reads the file into a data frame object.Any special characters or spaces in between will be replaced with " . "
:::

## ggplot - Histogram

```{r}
ggplot(data = realis,
       aes(x = `Unit Price ($ psm)`)) +
  geom_histogram()
```
